A Typed Programmatic Interface to Contracts on the Blockchain

### 2021-06-09 papers/execution-model/APLAS 2021/paper.tex
motivating example

%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{University of Freiburg, Germany \\
\email{\{doanha,thiemann\}@informatik.uni-freiburg.de}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  Smart contract applications on the blockchain can only reach their full potential if
  they can be integrated with traditional software systems. Invoking
  contracts and observing the state of the blockchain requires a
  programmatic interface. We propose a typed API for this
  purpose and establish some properties of the combined
  system. Specifically, we provide an execution model that
  enables us to prove type-safe interaction between programs and the 
  blockchain. We establish further properties of the model that
  give rise to requirements on the API. A prototype of the interface
  is implemented in OCaml for the Tezos blockchain.

\keywords{smart contract  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}
\label{sec:introduction}
First generation blockchains were primarily geared towards supporting
cryptocurrencies. The most prominent system of this kind is Bitcoin
\cite{bitcoin-whitepaper}. Though Bitcoin already features a
rudimentary programming language called Script, second generation
blockchains like the Ethereum blockchain \cite{eth-whitepaper} feature
Turing-complete programming facilities, called 
Smart Contracts. They provide robust trustworthy distributed computing
facilities even though the programs run on a peer-to-peer network with
untrusted participants. Each peer in the network runs the same program
and uses cryptographic methods to check the results among the other
peers and to create a persistent ledger of all transactions, the
blockchain, thus ensuring the integrity of the results. Third
generation blockchains, like Tezos \cite{tezos-whitepaper}, are even adaptable to
new requirements without breaking participating peers (no ``hard
forks'' required).

@@

Suppose you want to run an auction on the blockchain using a smart
contract. Once the contract is deployed, the
auction is open for bidding, but it should be closed after 48 hours. Such a timeout is a typical task for an oracle. Rather than
trusting a generic oracle, you prefer to roll your own mechanism that reliably closes the auction
after 48 hours have passed. So you want to write a program to do so.

As a bidder, you may want to implement a bidding
strategy that starts at a certain amount and increases the bid until a
limit is reached. Of course, you only want to increase your bid if
someone else placed a higher bid. So you also want to write a
program to implement this strategy.

Neither task can be implemented as a smart contract alone because both react
to external triggers. Closing require scheduling the invocation of a
contract at a certain time. Bidding 
requires watching the current highest bid of the contract and
react if another bidder places a higher bid.

@@

To avoid too much detail, we just outline the implementation of the entrypoints. The
\lstinline/%close/ entrypoint first checks its sender's address
against the owner's address in the store. Then it transfers the funds
to the owner, closes the contract by clearing the bidding flag, and
leaves it to the owner to deliver the
goods.\footnote{For simplicity we elide safeguarding by a third-party oracle.}
The \lstinline/%bid/ entrypoint immediately returns each bid that is not higher
than the existing highest bid. Otherwise, it keeps the funds
transferred, returns the previous highest bid to its owner, and stores
the current bidder as the new highest bidder.

Timed contract closing and strategic bidding cannot be implemented in
a smart contract, but could be implemented in separate programs that
interact with the blockchain. We present some idealized code for these tasks.

First the code that creates an auction contract and closes it after 48 hours.

\begin{lstlisting}[language=Caml,numbers=none,basicstyle=\small]
let _ = Cl.set_authority (my_address)
let initial_store = (true, (my_address, my_address))
let contract = Cl.originate(auction_script, 1, initial_store)
Time.sleep(48 * 3600);               (* 48h in seconds *)
Cl.call_contract(contract, left ()); (* entrypoint %close *)
\end{lstlisting}

We assume that \lstinline!Cl! is the module containing the contract
library, \lstinline/set_authority/ sets the sender address for
upcoming operations,  \lstinline/originate/ deploys its argument contract along
with the provided funds of one token and the \lstinline/initial_store/
\lstinline/auction_script/ is the contract from Listing~\ref{lst:auction-contract-header}, and 
\lstinline/call_contract/ invokes a contract with a parameter. The
\lstinline/initial_store/ sets the bidding flag, and both owner and
highest bidder to the sender account.

@@

\begin{lstlisting}[numbers=none,basicstyle=\small]
let rec poll (limit, step, ((bidding, _), (high_bid, _))) =
  if bidding and high_bid < limit then
    Cl.call_contract (contract, 
      right (min (high_bid + step, limit))); (* entrypoint %bid *)
    Time.sleep(5 * 60);
    let state = Cl.get_state (contract);
    poll (limit, step, state)
in
poll (limit, step, Cl.get_state (contract))
\end{lstlisting}

The \lstinline/get_state/ function obtains the storage of a contract from
the blockchain.

This code is idealized in several respects. Originating or running a contract
requires to propose a fee to the blockchain, which may or may not be
accepted.
Starting a contract may also time out for a variety of reasons. So just
invoking a contract with a fixed fee does not guarantee the contract's
execution.  Even if the invocation is locally accepted, it still takes
a couple of cycles before we can be sure the invocation is globally
accepted and incorporated in the blockchain.
Hence, after starting the invocation, we have to observe the fate of
this invocation. If it does not get incorporated, then we need to
analyse the reason and react accordingly. For example, if the
invocation was rejected because of an insufficient fee, we might want
to restart with an increased fee. Or we might decide to wait until the
invocation goes through without increasing the fee.

@@

In summary, a useful smart-contract-API  has facilities to
\begin{itemize}
\item query the current state of the blockchain (e.g., fees in the
  current block),
\item query storage and balance of a contract (to obtain
  the current highest bid),
\item originate contracts, invoke contracts, and initiate
  transfers. Hence, the API has to run on
  behalf of some account (by holding its private key).
\end{itemize}
All these facilities are supported by the (untyped) RPC interface of the Tezos
blockchain, which is the basis of our implementation.

>>>>

%
\authorrunning{~}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{University of Freiburg, Germany \\
\email{\{doanha,thiemann\}@informatik.uni-freiburg.de}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  Smart contract applications on the blockchain can only reach their full potential if
  they can be integrated with traditional software systems. Invoking
  contracts and observing the state of the blockchain requires a
  programmatic interface. We propose a typed API for this
  purpose and establish some properties of the combined
  system. Specifically, we provide an execution model that
  enables us to prove type-safe interaction between programs and the 
  blockchain. We establish further properties of the model that
  give rise to requirements on the API. A prototype of the interface
  is implemented in OCaml for the Tezos blockchain.
\end{abstract}

\keywords{smart contract  \and embedded domain specific language \and types.}
%
%
%
\section{Introduction}
\label{sec:introduction}
First generation blockchains were primarily geared towards supporting
cryptocurrencies. The most prominent system of this kind is Bitcoin
\cite{bitcoin-whitepaper}. Though Bitcoin already features a
rudimentary programming language called Script, second generation
blockchains like the Ethereum blockchain \cite{eth-whitepaper} feature
Turing-complete programming facilities, called 
Smart Contracts. They provide robust trustworthy distributed computing
facilities even though the programs run on a peer-to-peer network with
untrusted participants. Each peer in the network runs the same program
and uses cryptographic methods to check the results among the other
peers and to create a persistent ledger of all transactions, the
blockchain, thus ensuring the integrity of the results. Third
generation blockchains, like Tezos \cite{tezos-whitepaper}, are even adaptable to
new requirements without breaking participating peers (no ``hard
forks'' required).

@@

Suppose you want to implement a bidding strategy for an auction that
is deployed on the blockchain as a smart contract.
% Once the contract is deployed, the
% auction is open for bidding, but it should be closed after 48 hours. Such a timeout is a typical task for an oracle. Rather than
% trusting a generic oracle, you prefer to roll your own mechanism that reliably closes the auction
% after 48 hours have passed. So you want to write a program to do so. 
Your bidding strategy may start at a certain amount and increase the bid until a
limit is reached. Of course, you only want to increase your bid if
someone else placed a higher bid. So you want to write a
program to implement this strategy.

This task cannot be implemented as a smart contract without
cooperation of the auction contract because it reacts
on external triggers.
% Closing require scheduling the invocation of a
% contract at a certain time.
Bidding requires watching the current highest bid of the contract and
react if another bidder places a higher bid.
The auction contract could anticipate the need for such observations
by allowing bidders to register callbacks that are invoked when a
higher bid arrives. However, we cannot assume such cooperation of the
auction contract nor would we be willing to pay the the fee for running
that callback.

@@

To avoid too much detail, we only outline the implementation of the entrypoints. The
\lstinline/%close/ entrypoint first checks its sender's address
against the owner's address in the store. Then it transfers the funds
to the owner, closes the contract by clearing the bidding flag, and
leaves it to the owner to deliver the
goods.\footnote{For simplicity we elide safeguarding by a third-party oracle.}
The \lstinline/%bid/ entrypoint immediately returns each bid that is not higher
than the existing highest bid. Otherwise, it keeps the funds
transferred, returns the previous highest bid to its owner, and stores
the current bidder as the new highest bidder.

We present some idealized code for % timed contract closing and
strategic bidding as a separate program that interacts with the
blockchain.  The bidding strategy cannot be implemented as a smart
contract.
% A variation of timed contract closing could be
% implemented as Tezos contracts have access to the current time: A
% contract can enforce that bids arriving after a deadline are
% immediately rejected or that the auction cannot be closed before the
% deadline, but it cannot reliably execute the action associated with
% closing the contract at the deadline

In preparation, we need to obtain a typed handle for the contract. We
assume that \lstinline!Cl! is the module containing the contract
library. The respective library function
\lstinline!Cl.make_contract_from_hash!
takes the hash of the contract along with representations of
the types of the parameter and the storage (from module
\lstinline!Ct!). It checks the validity of the hash and the types with
the blockchain and returns a typed
handle, which is indexed with OCaml types corresponding to parameter
and storage type.
\begin{lstlisting}[language=Caml,numbers=none,basicstyle=\small]
# let auction = Cl.make_contract_hash auction_hash
#     ~parameter:(Ct.Or (Ct.Unit, Ct.Unit))
#     ~storage:(Ct.Pair (Ct.Bool, Ct.Pair (Ct.Addr, Ct.Addr))));;
val auction :
  ((unit, unit) Either.t,
   bool * (Cl.Addr.t * Cl.Addr.t)) Cl.contract
\end{lstlisting}

% First the code that creates an auction contract and closes it after 48 hours.

% \begin{lstlisting}[language=Caml,numbers=none,basicstyle=\small]
% let _ = Cl.set_authority (my_address)
% let initial_store = (true, (my_address, my_address))
% let contract = Cl.originate auction 1 initial_store
% Time.sleep(48 * 3600);               (* 48h in seconds *)
% Cl.call_contract contract (left ()); (* entrypoint %close *)
% \end{lstlisting}

% \lstinline/set_authority/ sets the sender address for
% upcoming operations,  \lstinline/originate/ deploys its argument contract along
% with the provided funds of one token and the \lstinline/initial_store/
% \lstinline/auction_script/ is the contract from Listing~\ref{lst:auction-contract-header}, and 
% \lstinline/call_contract/ invokes a contract with a parameter. The
% \lstinline/initial_store/ sets the bidding flag, and both owner and
% highest bidder to the sender account.

@@

\begin{lstlisting}[language=Caml,numbers=none,basicstyle=\small]
let rec poll limit step =
  let (bidding, (_, highest_bidder)) = Cl.get_storage auction;
  let high_bid = Cl.get_balance auction;
  if bidding && high_bid < limit then
    (if highest_bidder <> my_address then  (* entrypoint %bid *)
      try
        Cl.call_contract auction
          (right (min (high_bid + step, limit))) 
      with
      | Cl.FAILWITH message -> poll limit step;
    Time.sleep(5 * 60);
    poll limit step)
\end{lstlisting}

The functions \lstinline/get_storage/ and \lstinline/get_balance/
obtains the storage and current balance, respectively, of a contract from
the blockchain. They never fail. Function \lstinline/call_contract/
takes a typed handle and a parameter of suitable type. It indicates
failure by raising an exception. If failure is caused by the
\lstinline/FAILWITH/ instruction in the contract, then the
corresponding \lstinline/Cl.FAILWITH/ exception is raised, which
carries a string corresponding to the argument of the instruction. In
our particular example, the auction may fail with signaling the message
\lstinline/"closed"/ or \lstinline/"bid to low"/. Our
code ignores this message for simplicity.

This code is idealized in several respects. Originating or running a contract
requires to propose a fee to the blockchain, which may or may not be
accepted.
Starting a contract may also time out for a variety of reasons. So just
invoking a contract with a fixed fee does not guarantee the contract's
execution.  Even if the invocation is locally accepted, it still takes
a couple of cycles before we can be sure the invocation is globally
accepted and incorporated in the blockchain.
Hence, after starting the invocation, we have to observe the fate of
this invocation. If it does not get incorporated, then we need to
analyze the reason and react accordingly. For example, if the
invocation was rejected because of an insufficient fee, we might want
to restart with an increased fee. Or we might decide to wait until the
invocation goes through without increasing the fee.

@@

In summary, a useful smart-contract-API  has facilities to
\begin{itemize}
\item query the current state of the blockchain (e.g., fees in the
  current block),
\item query storage and balance of a contract (to obtain
  the current highest bid),
\item originate contracts, invoke contracts, and initiate
  transfers. Hence, the API has to run on
  behalf of some account (by holding its private key).
\end{itemize}
These facilities are supported by the (untyped) RPC interface of the Tezos
blockchain, which is the basis of our implementation.
